---
title: "\\textcolor{cobalt}{\\textbf{MSFS Coding Exercise}}"
author: "\\textcolor{cobalt}{\\textbf{Fernanda Yepez-Lopez}}"
output: 
  pdf_document:
    includes:
      in_header: preamble.tex
---

# Question
Among S&P 500 constituents as of October 1, 2024, do stocks that outperformed in Q4 2024 tend to continue outperforming in Q1 2025, or do they revert?
\vspace{1mm}

# Obtain the List of S&P 500 Constituents as of October 1st, 2024

I used the wayback machine website to look at the S&P 500 ETF Components https://www.slickcharts.com/sp500 on October 1st 2024. Then I copied the chart onto an Excel sheet 

\vspace{3mm}

```{r}
library(readxl)
library(quantmod)

sp500_Oct_data <- read_excel("S&P 500 Index Components.xlsx")
head(sp500_Oct_data)

tickers <- sp500_Oct_data$Symbol
head(tickers)
tickers <- gsub("\\.", "-", tickers)

```

# Adjusted Closing Prices 

### Helpful Dates Q4 2024

```{r}
q4_start <- as.Date("2024-10-01")
q4_end <- as.Date("2024-12-31")
q4_extended <-as.Date("2025-01-01")
```

### Helpful Dates Q1 2025

```{r}
q1_start <- as.Date("2025-01-01")
q1_end <- as.Date("2025-03-31")
q1_extended <-as.Date("2025-04-01")
```

### Write Function to Get Adjusted Closing Prices

We use adjusted price in order to  account for corporate actions such as stock splits, dividends, and rights offerings in the closing price. 

```{r}

find_adjusted <- function(tickers, start_date, end_date) {
  
   if (!is.vector(tickers) == 1) {
    tickers <- as.character(tickers)
  }

  price_list <- list()
  
  for (ticker in tickers) {
    
    tryCatch(
      {
        complete_ticker_data<- getSymbols(
          ticker, 
          src = "yahoo", 
          from = start_date, 
          to = end_date, 
          auto.assign = FALSE 
        )
        adj <-as.numeric(Ad(complete_ticker_data))
        price_list[[ticker]]=adj
      }, 
      error = function(e) {
        message(ticker)
      }
    )
  } 
  return(price_list)
}

```

### Adjusted Prices for all Ticker Q4
```{r}
q4_adjusted_prices <- find_adjusted(tickers, q4_start, q4_extended)
print(q4_adjusted_prices$AAPL)
```

### Adjusted Prices for all Ticker Q1

```{r}

q1_adjusted_prices <- find_adjusted(tickers,q1_start,q1_extended)

print(q1_adjusted_prices$AAPL)
```
### Adjusted Prices for Index Q4

```{r}
index <- "^GSPC"

q4_index_adj <- find_adjusted(index,q4_start,q4_extended)
head(q4_index_adj)
```
### Adjusted Prices for Index Q1

```{r}
q1_index_adj <- find_adjusted(index,q1_start,q1_extended)
head(q1_index_adj)

```

# Total Returns

### Function to Obtain Total Return of Each Ticker
```{r}
find_return <- function(adjusted_prices) {
  
  return_list <- list()
  
  for (ticker in names(adjusted_prices)) {
    tryCatch({
      start_adj <- as.numeric(adjusted_prices[[ticker]][1])
      end_adj <- as.numeric(tail(adjusted_prices[[ticker]], 1))
      return_val <- (end_adj - start_adj) / start_adj
      return_list[[ticker]] <- return_val
    }, error = function(e) {
      message(ticker)
    })
  }
  
  return(return_list)
}
```

### Total Return Q4
```{r}
q4_total_returns <-find_return(q4_adjusted_prices)
head(q4_total_returns)
```

### Total Return Q4
```{r}
q1_total_returns <- find_return(q1_adjusted_prices)
head(q1_total_returns)
```

### Index Return Q4
```{r}
q4_index_return <- find_return(q4_index_adj)
q4_index_return <- unlist(q4_index_return)
print(q4_index_return)
```

### Index Return Q1
```{r}
q1_index_return <- find_return(q1_index_adj)
q1_index_return <- unlist(q1_index_return)
print(q1_index_return)
```

# Outperforming and Underperforming Stock

### Sorting Q4 Stocks
```{r}

q4_outperforming_stocks <- c()
q4_underperforming_stocks <- c()

for (ticker in tickers) {
  if (q4_total_returns[[ticker]] > q4_index_return) {
    q4_outperforming_stocks <- c(q4_outperforming_stocks, ticker)
  }
  if (q4_total_returns[[ticker]] < q4_index_return) {
    q4_underperforming_stocks <- c(q4_underperforming_stocks, ticker)
  }
}

head(q4_outperforming_stocks)
head(q4_underperforming_stocks)

```

### Sorting Q1 Stocks
```{r}

q1_outperforming_stocks <- c()
q1_underperforming_stocks <- c()

for (ticker in tickers) {
  if (q1_total_returns[[ticker]] > q1_index_return) {
    q1_outperforming_stocks <- c(q1_outperforming_stocks, ticker)
  }
  if (q1_total_returns[[ticker]] < q1_index_return) {
    q1_underperforming_stocks <- c(q1_underperforming_stocks, ticker)
  }
}

head(q1_outperforming_stocks)
head(q1_underperforming_stocks)
```

### Verifying that all stocks are sorted Q4
```{r}
q4_total<-length(q4_outperforming_stocks)+length(q4_underperforming_stocks)
print(q4_total)

```

### Verifying that all stocks are sorted Q1
```{r}
q1_total<-length(q1_outperforming_stocks)+length(q1_underperforming_stocks)
print(q1_total)

```

# Contigency Table

### Sorting Q4
```{r}
Q4 <- list()

for(ticker in tickers){
  if(ticker %in% q4_outperforming_stocks){
    Q4[ticker] <- 1
  } else{
    Q4[ticker] <- 0
  }
}

#print(Q4)
```

### Sorting Q1
```{r}
Q1 <- list()

for(ticker in tickers){
  if(ticker %in% q1_outperforming_stocks){
    Q1[ticker] <- 1
  } else{
    Q1[ticker] <- 0
  }
}

#print(Q1)

```

### Dataframe
```{r}

df <- data.frame(
  ticker = names(Q4),
  Q4 = as.numeric(Q4),
  Q1 = as.numeric(Q1[names(Q4)])  
)

head(df)
```

### Contigency Table
```{r}
contingency_table <- table(Q4 = df$Q4, Q1 = df$Q1)

print(contingency_table)
```

### Probability Table
```{r}
prob_table <- prop.table(contingency_table, margin = 1)
rownames(prob_table) <- c("Underperform", "Outperform")
colnames(prob_table) <- c("Underperform", "Outperform")


prob_table <- round(prob_table, 3)
print(prob_table)
```
# Binomial Regression


### Regression

```{r}

performance <- glm(Q1~ Q4, data=df, family = binomial)

```

### Interpretations
```{r}
coeffs <- coef(performance)
intercept <- coeffs[1]
beta_Q4 <- coeffs[2]

p_q1_given_q4_under <- 1 / (1 + exp(-(intercept + beta_Q4 * 0)))
cat("P(Q1 Outperformer| Q4 Outperformer)=", p_q1_given_q4_under, "\n")

p_q1_given_q4_out <- 1 / (1 + exp(-(intercept + beta_Q4 * 1)))
cat("P(Q1 Outperformer| Q4 Underperformer)=", p_q1_given_q4_out)

```

# Visualizations 

```{r}
library(ggplot2)
prob_df <- as.data.frame(as.table(prob_table))
colnames(prob_df) <- c("Q4", "Q1", "Probability")


ggplot(prob_df, aes(x = Q4, y = Probability, fill = Q1)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = round(Probability, 3)),
            position = position_dodge(width = 0.7),
            vjust = -0.5, size = 4) +
  labs(
    title = "Probability of Q1 Performance Given Q4 Behavior",
    x = "Q4 Performance",
    y = "Probability"
  ) +
  scale_fill_manual(values = c("indianred", "steelblue")) +
  theme_minimal(base_size = 14)


```